\chapter{Appendix 7: Document Release History}


Date

Doc
Version

Status / Major Changes

10 Oct 1999

1.0 (Draft)

Protocol Version 1.0. First public draft release.

3 Dec 1999

1.0.1g

Protocol Version 1.0. Second public draft release. Much extended, incorporating packet format
layouts, APRS symbol tables, compressed data format, Mic-E format, telemetry format.

30 Apr 2000

1.0.1m

Protocol Version 1.0. Third public draft release.
Major additions/changes to the draft 1.0.1g specification:

\begin{itemize}
\item Added a section on Map Views and Range Scale.
\item 
\item   Changed Destination Address SSID description (specifying generic APRS digipeater paths) to apply to all packets, not just Mic-E packets.
\item Changed APRS destination “callsigns” to “destination addresses”.
\item Added TEL* to the list of generic destination addresses.
\item Added brief explanations of how several generic destination addresses are used.
\item Added “Grid-in-To-Address” (but marked as obsolete).
\item Extended the description of the Comment field, with pointers to what can appear in the field.
\item Added explanation of base 91.
\item Added paragraph on lack of consistency in on-air units, and default GPS datum = WGS84.
\item APRS Data Type Identifiers Table:
\item marked Shelter Data and Space Weather as reserved DTIs.
\item marked the - DTI as unused (previously erroneously allocated to Killed Objects).
\item marked the ' DTI to mean Current Mic-E data in Kenwood TM-D700 radios.
\item marked the ‘ DTI as not used in Kenwood TM-D700 radios.
\item Position Ambiguity: need only be specified in the latitude — the longitude will have the same
level of ambiguity.
\item Added the options of .../... and V V V /VV V V to express unknown course/speed.
\item Added DFS parameter table.
\item Added Quality table for BRG/NRQ data.
\item Position, DF and Compressed Report formats: split the format diagrams into two parts (with
and without timestamps).
\item DF Reports: added notes:
\end{itemize}

  \begin{itemize}
\item   BRG/NRQ data is only valid when the symbol is /\.
\item CSE=000 means the DF station is fixed, CSE non-zero means the station is moving.
\item Compressed position reports: corrected the multiplication/division constants for encoding/
decoding.
\item Mic-E chapter rewritten and expanded. Emphasized the need to ensure that non-printing
\item ASCII characters are not dropped. Corrected the Mic-E telemetry data format.
\item Expanded the introductory description of Objects/Items. All Objects must have a timestamp.
\item Added Area Object Extended Data field to Object and Item format diagrams.
\item Added Object/Item format diagrams with compressed location data.
\item Killed Objects/Items: now indicated by underscore after the name.
\item Re-categorized weather reports: Raw, Positionless and Complete.
\item Added a statement that temperatures below zero are expressed as -01 to -99.
\item Added the options of ... and V V V to express unknown weather parameter values.
\item Corrected the storm data format. Also, central pressure is now /ppppp (tenths of millibar).
\item Corrected the telemetry parameter data (now APRS messages instead of AX.25 UI beacons).
\item Added optional comment field to the Telemetry (T) format.
\item Added a section describing the handling of multiple message acknowledgements.
\item Added a section on NTS radiograms.
\item Added Bulletin/Announcement implementation recommendations.
\item Queries and Responses:
\item Query Names (e.g. APRSD): all upper-case.
\item A queried station need not respond if it has no relevant information to send.
\item A queried station should ignore any query type that it does not recognize.
\item APRSH: callsigns must be padded to 9 characters.
\item Added PING as a synonym of APRST.
\item Extended meteor scatter ERP beyond 810 watts, and added a lookup table.
\end{itemize}
\item   Maidenhead Locator: all letters must be transmitted in upper case, but may be received in
either upper or lower case.
\item Changed the definition of non-APRS packets — these are not APRS Status Messages, but
may optionally be treated as such.
\item APRS Symbols chapter substantially rewritten..
\item Added section on Symbol Precedence (where more than one symbol appears in an APRS
packet).
\item Clarified some of the descriptions in the APRS Symbol Tables.
\item Added overlay capability to the \a symbol (ARES/RACES etc).
\item Separated the 7-bit ASCII table from the Dec/Hex (0x80-0xff) conversion table.
\item Added several new entries and a units conversion table to the Glossary.
\item Added new references to NMEA sentence formats and Maidenhead Locator formats.
\end{itemize}

Date
29 Aug 2000

Doc
Version
1.0.1

Status / Major Changes
Protocol Version 1.0. Approved public release.


\section{Minor additions/changes to the draft 1.0.1m specification:}

\begin{itemize}
\item  Added Foreword.
\item  Replaced section on Map Views and Range Scale.
\item  APRS Software Version No: added APDxxx (Linux aprsd server).
\item  APRS Data Type Identifier: Designated [ as Maidenhead grid locator (but noted as obsolete).
\item  Position Ambiguity: added a bounding box example.
\item  Compressed Position Formats: for course/speed, corrected the range of possible values of the
“c” byte to 0–89.
\item  Mic-E: replaced the latitude example table, to show more explicitly how the N/S/E/W/Long
offset bits are encoded.
\item  Mic-E: removed the paragraph stating that there must be a space between the altitude and
comment text — no space is required.
\item  Mic-E: removed the note on inaccurate altitude data, as GPS Selective Availability has been
switched off.
\item  Object Reports: added timestamps to some of the examples (an Object Report must always
have a timestamp).
\item  Signposts: can be Objects or Items.
\item  Storm Data: changed central pressure format to /pppp (i.e. to the nearest millibar/hPascal).
\item  Storm Data: Hurricane Brenda examples: inserted a leading zero in the central pressure field
(central pressure is 4 digits).
\item  Telemetry Data: Added MIC as an alternative form of Sequence Number. MIC may or may
not be followed by a comma.
\item  Messages: added the reject message format.
\item  Appendix 1: Agrelo format: changed the separator between Bearing and Quality to /.
\item  Symbol Table: changed /( symbol from “Cloudy” to “Mobile Satellite Groundstation”.
\item  Reformatted the Units Conversion Table.

\end{itemize}

% END OF DOCUMENT
